### 2020-05-04 README.Rmd
Created README.Rmd

---
title: Using Google Community Mobility Reports to investigate the growth of COVID-19 in the United States
author:
  - name: Antonio Paez
    email: paezha@mcmaster.ca
    affiliation: McMaster University
    footnote: 1
address:
  - code: McMaster University
    address: School of Geography and Earth Sciences, McMaster University, Hamilton, ON, L8S 4K1, Canada
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: |
   In 2020 Google released a set of Community Mobility Reports (GCMR). These reports are based on the company's location-tracking capabilities and measure changes in mobility with respect to a baseline. This novel source of data offers an opportunity to investigate the way mobility potentially correlates with transmission of COVID-19. Using data from the New York Times on COVID-19 cases and GCMR, this paper presents an analysis of mobility levels and new daily reported cases of COVID-19 by state in the US. The results provide insights about the utility and interpretability of GCMR for COVID-19 research and decision-making.

journal: "Transportation Findings"
date: "`r Sys.Date()`"
bibliography: References.bib
#linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
---

@@

The main policy tool to control the spread of the COVID-19 pandemic has been stay-at-home orders. Concurrently, numerous efforts exist to track the progress and the impact of the pandemic, and new sources of data include the recently-released Google Community Mobility Reports (GCMR)\footnote{https://www.google.com/covid19/mobility/}, as well as The New York Times repository of COVID-19 data\footnote{https://github.com/nytimes/covid-19-data}. These two open data sets offer novel opportunities to investigate in quasi-real time the relationship between mobility patterns and transmission of COVID-19.

@@

```{r descriptive-statistics, echo=FALSE}
data.frame(Variable = c("New Cases",
                        "date",
                        "retail",
                        "groceries",
                        "parks",
                        "transit",
                        "work",
                        "residential"),
           Definition = c("New Daily Cases of COVID-19",
                        "Date",
                        "Retail and Recreation",
                        "Groceries and Pharmacies",
                        "Parks",
                        "Transit Stations",
                        "Workplaces",
                        "Residential"),
           min = c(as.character(min(c19$new_cases)),
                   as.character(min(c19$date)),
                   as.character(round(min(c19$retail), 2)),
                   as.character(round(min(c19$groceries), 2)),
                   as.character(round(min(c19$parks), 2)),
                   as.character(round(min(c19$transit), 2)),
                   as.character(round(min(c19$work), 2)),
                   as.character(round(min(c19$residential), 2))),
           median = c(as.character(median(c19$new_cases)),
                   as.character(median(c19$date)),
                   as.character(round(median(c19$retail), 2)),
                   as.character(round(median(c19$groceries), 2)),
                   as.character(round(median(c19$parks), 2)),
                   as.character(round(median(c19$transit), 2)),
                   as.character(round(median(c19$work), 2)),
                   as.character(round(median(c19$residential), 2))),
           max = c(as.character(max(c19$new_cases)),
                   as.character(max(c19$date)),
                   as.character(round(max(c19$retail), 2)),
                   as.character(round(max(c19$groceries), 2)),
                   as.character(round(max(c19$parks), 2)),
                   as.character(round(max(c19$transit), 2)),
                   as.character(round(max(c19$work), 2)),
                   as.character(round(max(c19$residential), 2))),
           sd = c(as.character(round(sd(c19$new_cases), 2)),
                   as.character(""),
                   as.character(round(sd(c19$retail), 2)),
                   as.character(round(sd(c19$groceries), 2)),
                   as.character(round(sd(c19$parks), 2)),
                   as.character(round(sd(c19$transit), 2)),
                   as.character(round(sd(c19$work), 2)),
                   as.character(round(sd(c19$residential), 2)))) %>%
  kable("latex",
        booktabs = TRUE,
        caption = "\\label{tab:descriptive-statistics}Descriptive statistics of the data set") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position")) %>%
  footnote(general = "All mobility indicators are lagged 11-day moving averages")
```

@@

```{r check-correlations, echo=FALSE}
#Check the correlation between the mobility indicators and incidence:
c19 %>% 
  select(log_new_cases, retail, groceries, parks, transit, work, residential) %>%
  cor(use = "pairwise.complete.obs") %>%
  kable("latex",
        booktabs = TRUE,
        digits = 2,
        caption = "\\label{tab:correlation-analysis}Simple correlation between log(New Cases) and the mobility indicators")  %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position")) %>%
  footnote(general = "All mobility indicators are lagged 11-day moving averages")
```

@@

```{r model-results, echo = FALSE}
data.frame(Variable = c("date",
                        "date^2",
                        "parks^2",
                        "parks",
                        "parks x residential",
                        "residential",
                        "residential^2",
                        "parks x date",
                        "residential x date",
                        "residential x date^2",
                        "NY",
                        "NY x date"),
           Coefficient = as.numeric(coefficients(mod1)),
           pval = as.numeric(summary(mod1)$coefficients[,4])) %>%
  mutate(pval = ifelse(pval >= 0.0001, round(pval, 4), "<0.0001")) %>%
  kable("latex",
        booktabs = TRUE,
        digits = 4,
        caption = "\\label{tab:model-results}Results of estimating regression model. Dependent variable is log(New Cases + 0.0001).",
        align = c("l", "c", "c"),
        col.names = c("Variable", "Coefficient Estimate", "p-value")) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  footnote(general = c(paste("Coefficient of Determination $R^2$=", round(summary(mod1)$r.squared, 3)),
                       paste("Adjusted Coefficient of Determination $R^2$=", round(summary(mod1)$adj.r.squared, 3)),
                       paste("Standard Error $\\\\sigma$=", round(summary(mod1)$sigma, 3))),
           escape = FALSE)
```

>>>>

---
output: github_document
always_allow_html: true
bibliography: References.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Using Google Community Mobility Reports to investigate the growth of COVID-19 in the United States

<!-- badges: start -->
<!-- badges: end -->

## Abstract

In 2020 Google released a set of Community Mobility Reports (GCMR). These reports are based on the company's location-tracking capabilities and measure changes in mobility with respect to a baseline. This novel source of data offers an opportunity to investigate the way mobility potentially correlates with transmission of COVID-19. Using data from the New York Times on COVID-19 cases and GCMR, this paper presents an analysis of mobility levels and new daily reported cases of COVID-19 by state in the US. The results provide insights about the utility and interpretability of GCMR for COVID-19 research and decision-making.

## Keywords

- COVID-19
- Google Community Mobility Reports
- Mobility
- Modelling
- Regression analysis
- Policy

@@

The main policy tool to control the spread of the COVID-19 pandemic has been stay-at-home orders. Concurrently, numerous efforts exist to track the progress and the impact of the pandemic, and new sources of data include the recently-released Google Community Mobility Reports [GCMR](https://www.google.com/covid19/mobility/), as well as The New York Times repository of [COVID-19 data](https://github.com/nytimes/covid-19-data). These two open data sets offer novel opportunities to investigate in quasi-real time the relationship between mobility patterns and transmission of COVID-19.

@@

```{r descriptive-statistics, echo=FALSE}
data.frame(Variable = c("New Cases",
                        "date",
                        "retail",
                        "groceries",
                        "parks",
                        "transit",
                        "work",
                        "residential"),
           Definition = c("New Daily Cases of COVID-19",
                        "Date",
                        "Retail and Recreation",
                        "Groceries and Pharmacies",
                        "Parks",
                        "Transit Stations",
                        "Workplaces",
                        "Residential"),
           min = c(as.character(min(c19$new_cases)),
                   as.character(min(c19$date)),
                   as.character(round(min(c19$retail), 2)),
                   as.character(round(min(c19$groceries), 2)),
                   as.character(round(min(c19$parks), 2)),
                   as.character(round(min(c19$transit), 2)),
                   as.character(round(min(c19$work), 2)),
                   as.character(round(min(c19$residential), 2))),
           median = c(as.character(median(c19$new_cases)),
                   as.character(median(c19$date)),
                   as.character(round(median(c19$retail), 2)),
                   as.character(round(median(c19$groceries), 2)),
                   as.character(round(median(c19$parks), 2)),
                   as.character(round(median(c19$transit), 2)),
                   as.character(round(median(c19$work), 2)),
                   as.character(round(median(c19$residential), 2))),
           max = c(as.character(max(c19$new_cases)),
                   as.character(max(c19$date)),
                   as.character(round(max(c19$retail), 2)),
                   as.character(round(max(c19$groceries), 2)),
                   as.character(round(max(c19$parks), 2)),
                   as.character(round(max(c19$transit), 2)),
                   as.character(round(max(c19$work), 2)),
                   as.character(round(max(c19$residential), 2))),
           sd = c(as.character(round(sd(c19$new_cases), 2)),
                   as.character(""),
                   as.character(round(sd(c19$retail), 2)),
                   as.character(round(sd(c19$groceries), 2)),
                   as.character(round(sd(c19$parks), 2)),
                   as.character(round(sd(c19$transit), 2)),
                   as.character(round(sd(c19$work), 2)),
                   as.character(round(sd(c19$residential), 2)))) %>%
  kable(#"latex",
        "html",
        booktabs = TRUE,
        caption = "\\label{tab:descriptive-statistics}Descriptive statistics of the data set") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position")) %>%
  footnote(general = "All mobility indicators are lagged 11-day moving averages")
```

@@

```{r check-correlations, echo=FALSE}
#Check the correlation between the mobility indicators and incidence:
c19 %>% 
  select(log_new_cases, retail, groceries, parks, transit, work, residential) %>%
  cor(use = "pairwise.complete.obs") %>%
  kable(#"latex",
        "html",
        booktabs = TRUE,
        digits = 2,
        caption = "\\label{tab:correlation-analysis}Simple correlation between log(New Cases) and the mobility indicators")  %>%
  #kable_styling(latex_options = c("striped", "scale_down", "hold_position")) %>%
  footnote(general = "All mobility indicators are lagged 11-day moving averages")
```

@@

```{r model-results, echo = FALSE}
data.frame(Variable = c("date",
                        "date^2",
                        "parks^2",
                        "parks",
                        "parks x residential",
                        "residential",
                        "residential^2",
                        "parks x date",
                        "residential x date",
                        "residential x date^2",
                        "NY",
                        "NY x date"),
           Coefficient = as.numeric(coefficients(mod1)),
           pval = as.numeric(summary(mod1)$coefficients[,4])) %>%
  mutate(pval = ifelse(pval >= 0.0001, round(pval, 4), "<0.0001")) %>%
  kable(#"latex",
        "html",
        booktabs = TRUE,
        digits = 4,
        caption = "\\label{tab:model-results}Results of estimating regression model. Dependent variable is log(New Cases + 0.0001).",
        align = c("l", "c", "c"),
        col.names = c("Variable", "Coefficient Estimate", "p-value")) %>%
  #kable_styling(latex_options = c("striped", "hold_position")) %>%
  footnote(general = c(paste("Coefficient of Determination $R^2$=", round(summary(mod1)$r.squared, 3)),
                       paste("Adjusted Coefficient of Determination $R^2$=", round(summary(mod1)$adj.r.squared, 3)),
                       paste("Standard Error $\\\\sigma$=", round(summary(mod1)$sigma, 3))),
           escape = FALSE)
```
